### 2021-11-21 cylindric-notes/Text.tex
more updates on RCA_omega

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{defin}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{fact}
\newtheorem{fact}{Fact}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\let\strokeL\L
\DeclareRobustCommand{\L}{\ifmmode\mathbf{L}\else\strokeL\fi}

\author{Daniel Rogozin}
\date{}
\title{Complete representability for canonical extensions of representable cylindric algebras}

\begin{document}
\maketitle

\section{The problem itself}

\begin{itemize}
\item Given $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? The conjecture is yes. \cite{hirsch2002relation}
\item Whether ${\bf RCA}_{\omega}$ is barely canonical. The conjecture is yes.
\end{itemize}
\section{Atomic Representations}

A representation of a Boolean algebra $\mathcal{B}$ is an embedding $h$ of $\mathcal{B}$ to some field of sets.

Let $a \in \mathcal{B}$ be an element of a Boolean algebra $\mathcal{B}$, $a$ is called an atom, if for every $b \in \mathcal{B}$
$b < a$ implies $b = 0$. That is, an atom is a minimal non-zero element. $\operatorname{At}(\mathcal{B})$ is the set of all atoms
of $\mathcal{B}$.

Let $\mathcal{B}$ be a Boolean algebra and $\mathcal{F}$ a field of sets such that $h : \mathcal{B} \to \mathcal{F}$ is a
representation of $\mathcal{B}$, then $\mathcal{B}$ is a complete representation of $\mathcal{B}$, if for every
$A \subseteq \mathcal{B}$ we have the following whenever $\Sigma A$ is defined:
\begin{center}
  $h(\Sigma A) = \bigcup h[A]$
\end{center}

A representation $h$ is called atomic, if $x \in h(1)$ there exists $b \in \operatorname{At}(\mathcal{B})$ such that $x \in h(b)$.

\begin{theorem} \label{completeboolean}
  Let $\mathcal{B}$ be a Boolean algebra, then $\mathcal{B}$ is atomic iff $\mathcal{B}$ is completely representable. See \cite[Corollary 6]{hirsch1997complete}.
\end{theorem}

\section{BAOs and Duality}

By default, we assume that all operators are at most unary. Here is the rigorous definition:

\begin{defin}

$ $

  \begin{enumerate}
    \item Let $\mathcal{B} = \langle B, +, -, 0, 1 \rangle$ be a Boolean algebra. An operator is a function $\Omega : B \to B$ satisfying the following conditions:
    \begin{itemize}
      \item Normality: $\Omega(0) = 0$
      \item Additivity: $\Omega(b + b') = \Omega(b) + \Omega(b')$
    \end{itemize}
    \item Let $I$ be an index set, a Boolean algebra with operators (BAO) is an algebra $\langle B, +, -, 0, 1, (\Omega_{i})_{i \in I} \rangle$ such that $\langle B, +, -, 0, 1 \rangle$ is a Boolean algebra and for each $i \in I$ $\Omega_{i}$ is an operator.
  \end{enumerate}
\end{defin}

\begin{defin} Let $\mathcal{B} = \langle B, +, -, 0, 1, (\Omega_{i})_{i \in I} \rangle$ be a BAO, then

  \begin{enumerate}
    \item An opetator $\Omega$ is completely additive, if for every $X \subseteq B$ such that $\Sigma X$ is defined, one has

    \begin{center}
      $\Omega(\sum X) = \sum \limits_{x \in X} \Omega(x)$
    \end{center}
    \item $\mathcal{B}$ is completely additive, if for each $i \in I$ $\Omega_{i}$ is additive,
    \item A class $\mathcal{K}$ of BAOs is completely additive, if every $\mathcal{B} \in \mathcal{K}$ is completely additive.
  \end{enumerate}
\end{defin}

\subsection{Atom structures and canonical extensions}

\begin{defin}  Let $I$ be an index set and $(\Omega_i)_{i \in I}$ a set of function symbols
\begin{enumerate}
  \item A structure is a relational structrure
  $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$
  such that $R_{i}$ is a binary relation symbol for a function symbol $\Omega_{i \in I}$ with the corresponding index,
  \item Let $\mathcal{B}$ be an atomic BAO of the signature $I$,
the atom structure of $\mathcal{B}$, written as $\mathfrak{At} \mathcal{B}$, is a structure $\langle \operatorname{At}(\mathcal{B}), (R_{i})_{i \in I} \rangle$ such that for all
$a, b \in \operatorname{At}(\mathcal{B})$ and for all $i \in I$
\begin{center}
  $\mathfrak{At} \mathcal{B} \models R_{i}(a,b)$ iff $\mathcal{B} \models a \leq \Omega_{i}(b)$
\end{center}
\item Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ be an atom structure, the complex algebra of $\mathcal{F}$, written as $\mathfrak{Cm} \mathcal{F}$, is a BAO
$\langle \mathcal{P}(W), \cup, -, \emptyset, W, (\Omega_{R_{i}})_{i \in I} \rangle$ such that
for all $X \subseteq W$ and for each $i \in I$:
\begin{center}
  $\Omega_{R_{i}}(X) = \{ a \in W \: | \: \exists b \in X \: \mathcal{F} \models R_{i}(a, b)\}$
\end{center}
\end{enumerate}
\end{defin}

\begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ and $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, then a function $f : \mathcal{F} \to \mathcal{F}'$ is a bounded morphism, if the following holds:
\begin{enumerate}
\item $x R_i y$ implies $f(x) {R'}_i f(y)$;
\item $f(x) {R'}_i z$, then there exists $y \in W$ such that $x R_i y$ and $f(y) = z$.
\end{enumerate}
A bounded morphism $f : \mathcal{F} \to \mathcal{F}'$ is a $p$-morphism, if $f$ is onto. $\mathcal{F} \twoheadrightarrow \mathcal{F}'$ iff there exists a $p$-morphism from $\mathcal{F}$ onto $\mathcal{F}'$, or $\mathcal{F}'$ is a $p$-morphic image of $\mathcal{F}$.
\end{defin}

\begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ is an inner substructure \footnote{Or alternatively, a generated subframe} of $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, if $W \subseteq W'$ and the embedding $\mathcal{F} \hookrightarrow \mathcal{F}'$ is a bounded morphism. Let $\mathbb{F}$ be a class atom structures, then $\mathbb{S}(\mathbb{F})$ is the closure of $\mathbb{F}$ under generated subframes.
\end{defin}

Let $\mathbb{F}$ be a class of structures, define:
\begin{enumerate}
\item $\mathfrak{Cm}(\mathbb{F}) = \{ \mathcal{B} \: | \: \mathcal{B} \cong \mathfrak{Cm}(\mathcal{F}) \text{ for some $\mathcal{F} \in {\bf F}$}\}$.
\item ${\bf Up}(\mathbb{F})$ is the class of structures isomorphic to disjoint unions of elements of $\mathbb{F}$.
\item ${\bf S}(\mathbb{F})$ is the closure of $\mathbb{F}$ under inner substructures.
\end{enumerate}

Let $A$ be a non-empty subset of a Boolean algebra $\mathcal{B}$, $A$ is a \emph{filter}, if $A$ is closed under finite infima and it is upward closed. $A$ is an ultrafilter, if it has no non-trivial extensions. That is, if $A \subseteq A'$, then $A' = \mathcal{B}$. This is a well-known fact that every filter can be extended to a maximal one using Zorn's lemma.

The following definition is due to, for example, \cite[Definition 5.40]{venema2010}.

\begin{defin}
  Let $\mathcal{B} = \langle B, +, -, 0, 1, (\Omega_i)_{i \in I} \rangle$ be a BAO and ${\bf Spec}(\mathcal{B})$ the set of its ultrafilters. The ultrafilter frame of $\mathcal{B}$ (or the canonical frame) is a relational structure $\mathcal{F}_{\mathcal{B}} = \langle {\bf Spec}(\mathcal{B}), R_{\Omega_i} \rangle$ such that for all ultrafilters $U_1, U_2$ one has
  \begin{center}
    ${\bf Spec}(\mathcal{B}) \models R_{\Omega_i}(U_1, U_2)$ iff $\{ \Omega_i(b) \: | \: b \in U_1 \} \subseteq U_2$.
  \end{center}
\end{defin}

Given $\mathcal{B}$ be a BAO, we denoted as $\mathcal{B}^{+}$ as the complex algebra of the canonical frame $\mathfrak{Cm}(\mathcal{F}_{\mathcal{B}})$, that is, \emph{the canonical extension} of $\mathcal{B}$.
A class of BAOs ${\bf K}$ is canonical, if it is closed under canonical extensions. That is, $\mathcal{B}^{+} \in {\bf K}$ whenever $\mathcal{B} \in {\bf K}$.

\begin{theorem} Let $\mathcal{A}$, $\mathcal{B}$ be BAOs,

\begin{enumerate}
  \item There exists $\iota : \mathcal{A} \hookrightarrow \mathcal{A}^{+}$ such that
  $\iota : a \mapsto \{ \gamma \in {\bf Spec}(\mathcal{A}) \: | \: a \in \gamma \}$.
  \item $i : \mathcal{A} \hookrightarrow \mathcal{B}$ implies
  $i^{+} : \mathcal{A}^{+} \hookrightarrow \mathcal{B}^{+}$
\end{enumerate}
\end{theorem}

\section{Representable cylindric algebras}

Let $\alpha$ be an ordinal. Denote $\{ f \: | \: f \alpha \to U\}$ as $\prescript{\alpha}{}U$. $x_i$ stands for $x(i)$, where
$x \in \prescript{\alpha}{}U$ and $i < \alpha$.

A subset of $\prescript{\alpha}{}U$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the \emph{$i,j$-diagonal} $D_{ij}$ is the set of all elements of $\prescript{\alpha}{}U$ such that $y_i = y_j$.

If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except, perhaps, the $i$-th one. To be more precise,
\begin{center}
$C_i X = \{ y \in \prescript{\alpha}{}U \: |
\: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
\end{center}
We define the following equivalence relation for $i < \alpha$ and $x, y \in \prescript{\alpha}{}U$:
\begin{center}
 $x \equiv_i y \Leftrightarrow \forall j \in \alpha \: (i \neq j \Rightarrow x(i) = y(j))$
\end{center}
Then one may reformulate the definition of the $i$-th cylindrification in the following way:

\begin{center}
 $C_i X = \{ y \in \prescript{\alpha}{}U \: | \: \exists x \in X \:\: x \equiv_i y \}$
\end{center}

According to this version of the definiton, one may think of the cylindrification as an ${\bf S}5$ modal operator.

\begin{defin}
 A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
   with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $(D_{ij})_{i, j < \alpha}$, the cylindrifications $(C_i)_{i < \alpha}$. A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$.
   ${\bf Cs}_{\alpha}$ denotes the class of all cylindric set algebras of dimension $\alpha$.
\end{defin}

\begin{defin}
   A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
   \begin{itemize}
     \item $\mathcal{B}$ is a Boolean algebra, for each $i, j < \alpha$ $c_i$ is an operator and $d_{ij} \in \mathcal{B}$
     \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \cdot c_i b) = c_i a \cdot c_i b$ and $d_{ii} = 1$
     \item For every $i, j < \alpha$, $c_i c_j a = c_j c_i a$
     \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \cdot d_{jk})$
     \item If $i \neq j$, then $c_i (d_{ij} \cdot a) \cdot c_i (d_{ij} \cdot - a) = 0$
   \end{itemize}
   ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$.
\end{defin}

One may define a representation of a cylindric algebra explicitly in the following way:

\begin{defin}
 Let $\mathcal{A}$ be a cylindric algebra of dimension $\alpha$. A representation of $\mathcal{A}$ over the non-empty domain $X$ is a map $f : \mathcal{A} \hookrightarrow 2^{\prescript{\alpha}{}U}$ such that:
 \begin{enumerate}
   \item $f(1) = \bigcup \limits_{i \in I} \prescript{\alpha}{}X_i$ for some disjoint family $\{X_i\}_{i \in I}$ where each $X_i \subseteq X$
   \item $h : \mathcal{A} \to 2^{f(1)}$ is a representation of a Boolean reduct
   \item for all $\lambda, \eta < \alpha$, $x \in h(d_{\lambda \eta})$ iff $x_{\lambda} = x_{\eta}$
   \item for all $\lambda < \alpha$ and $a \in \mathcal{A}$, $x \in h(c_{\lambda}(a))$ iff there is $y \in X$ such that $x[\lambda \mapsto y] \in h(a)$
 \end{enumerate}
\end{defin}

An $\alpha$-dimensional cylindric algebra $C$ is representable, if there exists a representation of $h$.
${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$. In particular, we are interested in the case $\alpha = \omega$.

It is well known that ${\bf RCA}_{\alpha}$ is a variety, ${\bf RCA}_{\alpha}$ ($\alpha \leq 2$) is finitely axiomatisable and ${\bf RCA}_{\alpha}$ ($2 < \alpha < \omega$) has no finite axiomatisation, see \cite{Henkin1988-HENCAP-4}.

Let $\mathcal{A} \in {\bf CA}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if its representation preserves all existing suprema. In other words, $\mathcal{A}$ is \emph{completely representable}.

\section{${\bf RCA}_{\omega}$ and canonicity}

The following definition of an $\omega$-frame is due to \cite{Venema2013}.
\begin{defin}
  A cylindric $\omega$-frame is a structure $\mathcal{F} = \langle W, (R_i)_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ where $(R_i)_{i < \omega}$ are binary relations and $(E_ij)_{i, j < \omega}$ are unary relations such that, for all $i, j, k < \omega$:
  \begin{enumerate}
  \item Every $R_i$ is an equivalence relation on $W$,
  \item $R_i \circ R_j = R_j \circ R_i$, that is, the set $(R_i)_{i < \omega}$ forms a commutative semigroup under composition.
  \item For all $x \in W$, $E_{ii}(x)$ holds.
  \item For all $x, y, z \in W$, $x R_i y \: \& \: E_{ij}(y) \: \& \: x R_i z \: \& \: E_{ij}(y)$ implies $y = z$.
  \item For all $x \in W$, $E_{ij}(x)$ iff there exists $y \in W$ such that $x R_k y$, $E_{ik}(y)$, and $E_{kj}(y)$.
  \end{enumerate}
  $\mathcal{C}\mathfrak{a}_{\omega}$ is the class of all $\omega$-frames.

  If $\mathcal{F} \in \mathcal{C}\mathfrak{a}_{\omega}$ and $x \in \mathcal{F}$, then $\mathcal{F}^{x}$ is a generated subframe generated by $x$, which is defined standardly. Generally, $\mathcal{F}_1$ is a generated subframe of $\mathcal{F}_2$, if $\underline{\mathcal{F}_1} \subseteq \underline{\mathcal{F}_2}$ and $\underline{\mathcal{F}_1}$ is closed under under ${R_i}_2$ equivalences for every $i < \omega$. That is:
  \begin{center}
  For all $i < \omega$ and $x \in \mathcal{F}_1$, we have ${R_i}_2(x) \subseteq \mathcal{F}_1$ and, thus, ${R_i}_1(x) = {R_i}_2(x)$.
  \end{center}
\end{defin}

We have the following connection betweens $\omega$-frames and their generated subframes, which is standard for modal logic:
\begin{prop}
Let $\mathcal{F} \in \mathcal{C}\mathfrak{a}_{\omega}$, then
\begin{enumerate}
\item $\mathcal{F} = \coprod \limits_{x \in \mathcal{F}} \mathcal{F}^{x}$,
\item $\mathfrak{Cm}(\mathcal{F}) \cong \prod \limits_{x \in \mathcal{F}} \mathfrak{Cm}(\mathcal{F}^{x})$,
\item $\mathfrak{Cm}(\mathcal{F}^{x})$ is subdirectly irreducible.
\end{enumerate}
\end{prop}

It is known that $\mathcal{C}\mathfrak{a}_{\omega}$ forms an elementary class, since one can express the conditions of an $\omega$-frame with the first-order language.

The following fact is by Venema, see \cite[Proposition 2.1.5]{Venema2013}:
\begin{prop}
An $\omega$-frame $\mathcal{F}$ is cylindric iff $\mathfrak{Cm}(\mathcal{F})$ is a cylindric algebra of dimension $\omega$.
\end{prop}
A cylindric $\omega$-frame $\mathcal{F}$ is completely representable, if $\mathfrak{Cm}(\mathcal{F})$ is completely representable as a cylindric algebra of dimension $\omega$.

We are interested in the special case of cylindric $\omega$-frames called Cartesian structure of dimension $\omega$. To be more precise:

\begin{defin}
Let $U$ be a set and $V \subseteq \prescript{\omega}{}U$ be a non-empty subset of the full Cartesian space of dimension $\omega$, then an $\alpha$-dimension Cartesian structure generated by $V$ is an $\omega$-frame $\mathfrak{S}(V) = \langle V, (R_{i})_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ such that:
\begin{enumerate}
\item $R_i = \{ (w, v) \: | \: w, v \in V, w_k = w_k, k < \omega, i \neq k \}$
\item $E_{ij} = \{ w \in V \: | \: w_i = w_j \}$
\end{enumerate}
$\mathfrak{S}(\prescript{\omega}{}U)$ is the full $\omega$-dimensional Cartesian structure. $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$ is the class of all full $\omega$-dimensional Cartesian structures.
\end{defin}

Clearly $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega} \subseteq \mathcal{C}\mathfrak{a}_{\omega}$.

We have the following connection between ${\bf RCA}_{\omega}$, ${\bf IGs}_{\omega}$, and complex algebras of full Cartesian structures:
\begin{center}
${\bf RCA}_{\omega} = {\bf IGs}_{\omega} = {\bf S} \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega} = {\bf S P} \mathfrak{Cm}  \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$.
\end{center}
This follows from the fact that ${\bf Cs}_{\omega} = \mathfrak{Cm} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$. Every generalised cylindric set algebra is a subdirect product of cylindric set algebras, thus, a generalised cylindric set algebra is a complex algebra of disjoint union of some full Cartesian spaces. But ${\bf RCA}_{\omega}$ is the closure of ${\bf Cs}_{\omega}$ under isomorphism.

\begin{defin}
The weak Cartesian space with base $U$ and dimension $\omega$ determined by $x \in \prescript{\omega}{}U$ is the set:
\begin{center}
$\prescript{\omega}{}U^{(x)} = \{ y \in \prescript{\omega}{}U \: | \: |\{ k < \omega \: | \: x_k \neq y_k \}| < \aleph_0 \}$
\end{center}
$\mathfrak{S}(\prescript{\omega}{}U^{(x)})$ is a weak Cartesian structure of dimension $\omega$. $\mathcal{W} \mathfrak{ct}_{\omega}$ is the class of all weak Cartesian structure of dimension $\omega$ up to isomorphism.
\end{defin}
Note that we have $\mathcal{W} \mathfrak{ct}_{\omega} \subseteq \mathcal{C}\mathfrak{a}_{\omega}$.

Every cylindric set algebra is a subalgebra of some complex algebra induced by an $\omega$-dimensional Cartesian structure. In other words,

\begin{lemma}
${\bf I}{\bf Cs}_{\omega} = {\bf S} \mathfrak{Cm} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$.
\end{lemma}

In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$. The following results are due to Goldblatt \cite{goldblatt1995elementary}. This denotes that a cylindric algebra of dimension algebra is representable iff it is isomorphic to a subalgebra of the complex algebra of disjoint sum of some full $\omega$-dimensional Cartesian structure. Assuming the duality, this is equivalent to the standard definition of representability formulated in terms of sublagebras of subdirect products.

\begin{lemma} \label{rcachar}
${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{W} \mathfrak{ct}_{\omega} = {\bf IGws}_{\omega}$
\end{lemma}

\begin{proof}

\end{proof}

Here we use the following fact related to canonical varieties generated by some class of complex algebras. Let ${\bf K}$ be an elementary class of relational structures, then:
\begin{center}
If ${\bf K}$ is closed under $p$-morphic images, generated subframes, and disjoint unious, \\ then ${\bf S} \mathfrak{Cm} {\bf K}$ is a canonical variety.
\end{center}
One may think of this fact a more abstract version of Fine's theorem which claims that every elementary modal logic is canonical \cite{fine1975some}. This version denotes the same fact, but it is formulated in terms of varieties BAOs generated by complex algebras of some atom structures.

This is a specialised version of \cite[Theorem 4.4]{goldblatt1995elementary} formulated for dimension $\omega$.

\begin{theorem}
${\bf RCA}_{\omega}$ is a canonical variety.
\end{theorem}

\begin{proof}
By Lemma~\ref{rcachar},
\end{proof}

\section{Representability games}

\bibliographystyle{plain}
\bibliography{Text}

\end{document}