### 2023-02-06 residuated-semigroups/Text.tex
lambek and residuated stuff

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{definition}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{claim}
\newtheorem{claim}{Claim}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{fact}
\newtheorem{fact}{Fact}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\let\strokeL\L
\DeclareRobustCommand{\L}{\ifmmode\mathbf{L}\else\strokeL\fi}

\author{Daniel Rogozin}
\date{}
\title{Finite model property for residuated semigroups and related remarks}

\begin{document}
\maketitle

\nocite{*}

\section{Finite networks for atomic formulas}

Let $PV = \{ p_i \: | \: i < \omega \}$ be the set of propositional variables (or atomic types).
The set of formulas is generated by the following grammar:
\begin{center}
$\varphi, \psi ::= p \: | \: (\varphi \bullet \psi) \: | \: (\varphi \backslash \psi) \: | \: (\varphi / \psi)$
\end{center}

The Lambek calculus is defined as a Gentzen-style sequent calculus:

\begin{prooftree}
\AxiomC{$ $}
\RightLabel{\scriptsize{ax}}
\UnaryInfC{$p \rightarrow p$}
\end{prooftree}

\begin{minipage}{0.5\textwidth}
\begin{flushleft}
      \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow \varphi$}
    \AxiomC{$\Delta, \psi, \Theta \rightarrow \theta$}
    \RightLabel{$\backslash \rightarrow$}
    \BinaryInfC{$\Delta, \Gamma, \varphi \backslash \psi, \Theta \rightarrow \theta$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow \varphi$}
    \AxiomC{$\Delta, \psi, \Theta \rightarrow \theta$}
    \RightLabel{$/ \rightarrow$}
    \BinaryInfC{$\Delta, \psi / \varphi, \Gamma, \Theta \rightarrow \theta$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma, \varphi, \psi, \Delta \rightarrow \theta$}
    \RightLabel{$\bullet \rightarrow$}
    \UnaryInfC{$\Gamma, \varphi \bullet \psi, \Delta \rightarrow \theta$}
  \end{prooftree}
\end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}
  \begin{prooftree}
    \AxiomC{$\varphi, \Pi \rightarrow \psi$}
    \RightLabel{$\rightarrow \backslash, \Pi \text{ is non-empty}$}
    \UnaryInfC{$\Pi \rightarrow \varphi \backslash \psi$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Pi, \varphi \rightarrow \psi$}
    \RightLabel{$\rightarrow /, \Pi \text{ is non-empty}$}
    \UnaryInfC{$\Pi \rightarrow \psi / \varphi$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow \varphi$}
    \AxiomC{$\Delta \rightarrow \varphi$}
    \RightLabel{$\rightarrow \bullet$}
    \BinaryInfC{$\Gamma, \Delta \rightarrow \varphi \bullet \psi$}
  \end{prooftree}
\end{flushright}
\end{minipage}

\begin{theorem}
Let $RS$ be the class of all residuated semigroups, then $\Gamma \rightarrow \varphi$ iff $RS \models \Gamma \rightarrow \varphi$
\end{theorem}

\section{Representability networks for at most countable residuated semigroups}

Let $\mathcal{A}$ be a residuated semigroup, an $\mathcal{A}$-network is a triple $\mathcal{N} = (V, E, l)$, where
\begin{itemize}
\item $(V, E)$ is a directed graph with no loops and $E$ is transitive
\item $l : E \to \mathcal{A}$ such that
\begin{itemize}
\item $l(x, z) \leq l(x, y) ; l(y,z)$, whenever $(x, y), (y, z) \in E$, for all $x, y, z \in V$.
\item 
\end{itemize}
\end{itemize}

Let $n \leq \omega$

\section{Games for the FMP}

\bibliographystyle{alpha}
\bibliography{Text}
\end{document}